\section{NP-Hardness}
\begin{definition}
\emph{P} is the set of decision problems that can be solved in polynomial time. Intuitively, P is the set of problems that can be solved quickly.
\end{definition}
\begin{definition}
\emph{NP} is the set of decision problems with the following property: if the answer is \Var{Yes}, then there is a \emph{proof} of this fact that can be checked in polynomial time. Intuitively, NP is the set of decision problems where we can verify a \Var{Yes} answer quickly if we have the solution in front of us.
\end{definition}
\begin{definition}
\emph{co-NP} is essentially the opposite of NP. If the answer to a problem in co-NP is \Var{No}, then there is a proof of this fact that can be checked in polynomial time.
\end{definition}
Every decision problem in P is also in NP and also in co-NP.
\begin{definition}
A problem $\Pi$ is \emph{NP-hard} if a polynomial-time algorithm for $\Pi$ would imply a polynomial-time algorithm for every problem in NP.
\end{definition}
\begin{definition}
A problem $\Pi$ is \emph{NP-complete} if it is both NP-hard and an element of NP.
\end{definition}
\begin{theorem}[Cook-Levin Theorem]
Circuit satisfiability is NP-complete.
\end{theorem}
To prove that problem $A$ is NP-hard, reduce a known NP-hard problem to $A$.
\begin{definition}
A \emph{many-one} reduction from one language $L' \subseteq \Sigma^*$ is a function $f : \Sigma^* \rightarrow \Sigma^*$ such that $x \in L'$ iff $f(x) \in L$. A \emph{language} $L$ is NP-hard iff, for any language $L' \in NP$, there is a \emph{many-one} reduction from $L'$ to $L$ that can be computed in polynomial time.
\end{definition}
\subsection{NP-Hard Problems}
\begin{itemize}
	\item SAT
	\item 3SAT
	\item Maximum Independent Set: find the size of the largest subset of the vertices of a graph with no edges between them
	\item Clique: Compute the number of nodes in its largest complete subgraph
	\item Vertex Cover: Smallest set of vertices that touch every edge in the graph
	\item Graph Coloring: Find the smallest possible number of colors in a legal coloring such that every edge has two different colors at its endpoints
	\item Hamiltonian Cycle: find a cycle that visits each vertex in a graph exactly once
	\item Subset Sum: Given a set $X$ of positive integers and an integer $t$, determine whether $X$ has a subset whose elements sum to $t$
	\item Planar Circuit SAT: Given a boolean circuit that can be embedded in the plane so that no two wires cross, is there an input that makes the circuit output \Var{True}
	\item Not All Equal 3SAT: Given a 3CNF formula, is there an assignment of values to the variables so that every clause contains at least one \Var{True} literal \emph{and} at least one \Var{False} literal?
	\item Exact 3-Dimensional Matching: Given a set $S$ and a collection of three-element subsets of $S$, called \emph{triples}, is there a sub-collection of disjoint triples that exactly cover $S$?
	\item Partition: Given a set $S$ of $n$ integers, are there subsets $A$ and $B$ such that $A \cup B = S$, $A \cap B = \emptyset$, and $\sum_{a \in A} a = \sum_{b \in B} b$?
	\item 3Partition: Given a set $S$ of $3n$ integers, can it be partitioned into $n$ disjoint three-element subsets, such that every subset has exactly the same sum?
	\item Set Cover: Given a collection of sets $\mathscr{S} = \{ S_1, S_2, \ldots, S_m \}$, find the smallest sub-collection of $S_i$'s that contains all the elements of $\bigcup_i S_i$
	\item Hitting Set: Given a collection of sets $\mathscr{S} = \{ S_1, S_2, \ldots, S_m \}$, find the minimum number of elements of $\bigcup_i S_i$ that hit every set in $\mathscr{S}$
	\item Hamiltonian Path: Given a graph $G$, is there a path in $G$ that visits every vertex exactly once?
	\item Longest Path: Given a non-negatively weighted graph $G$ and two vertices $u$ and $v$, what is the longest simple path from $u$ to $v$ in the graph? A path is \emph{simple} if it visits each vertex at most once.
	\item Steiner Tree: Given a weighted, undirected graph $G$ with some of the vertices marked, what is the minimum-weight subtree of $G$ that contains every marked vertex?
\end{itemize}