\section{SAT and CNF-SAT}

\subsection{Formula Satisfiability/SAT}
Given a boolean formula like $(a \vee b \vee c \vee \overline{d}) \Leftrightarrow ((b \wedge \overline{c}) \vee \overline{ ( \overline{a} \Rightarrow d) } \vee (c \neq a \wedge b))$, is it possible to assign boolean values to $a, b, c, \ldots$ so that the formula evaluates to \Var{True}?\\

We can transform any boolean circuit to a formula in linear time (using depth-first search), and the size of the resulting formula is only a constant factor larger than the size of the circuit. Thus, we have a polynomial-time reduction from circuit satisfiability to SAT. Thus, SAT is NP-hard.\\

To prove that a boolean formula is satisfiable, we only have to specify an assignment to the variables that makes the formula \Var{True}. We can check the proof in linear time just by reading the formula from left to right, evaluating as we go. So SAT is also in NP, and thus is NP-complete.

\subsection{CNF}
A boolean form is in \emph{conjunctive normal form} if it is a conjunction (\alg{and}) of several clauses, each of which is the disjunction (\alg{or}) of several \emph{literals}, each of which is either a variable or its negation. For example, $(a \vee b \vee c \vee d) \wedge (b \vee \overline{c} \vee \overline{d}) \wedge (\overline{a} \vee c \vee d) \wedge (a \vee \overline{b})$.

\subsection{3CNF}
A 3CNF formula is a CNF formula with exactly 3 literals per clause; the previous example is not a 3CNF formula, since its first clause has four literals and its last clause has only two. 3SAT is just SAT restricted to 3CNF formulas: Given a 3CNF formula, is there an assignment to the variables that makes the formula evaluate to \Var{True}?

\begin{enumerate}
	\item Make sure every \alg{and} and \alg{or} gate has only two inputs. If any gate has $k > 2$ inputs, replace it with a binary tree of $k - 1$ two-input gates.
	\item Write down the circuit as a formula, with one clause per gate.
	\item Change every gate clause into a CNF formula
	\item Make sure every clause has exactly three literals
\end{enumerate}

Every binary gate in the original circuit will be transformed into at most 5 clauses.